Avoiding the Global Sort: \break A Faster Contour Tree Algorithm \footnote{The full updated version of this paper is available on the arXiv \cite{rs-mvrms-14}}

### 2016-08-01 paint-contour-tree.tex
typos fixed

We prove an important lemma. The cost of heap operations corresponding 
to a buffered sub-shrub to be charged to tall support chains.

@@

There is a grand total of $100 \sum_{p \in \buff} |\loss(p)|$ tokens, which
(by the buffered property of $\buff$) is strictly more than $\sum_{p \in \buff} |p|$.
We will place all these tokens on the nodes in $\buff$, using the following process.
We reserve topologically sort (descendants before ancestors) the nodes in $\buff$.
For each $p$ in order, we will place tokens in $T_p$ on the nodes of the subtree $\buff_p$.
Take an arbitrary token in $T_p$, and find the smallest (in the ordering) descendant $q$
of $p$ that currently has strictly less than $|q|$ tokens. Place the token on $q$,
and repeat. This is repeated until all descendants $q$ have $|q|$ tokens, or
all tokens in $T_p$ are used up. When a node $q$ has $|q|$ tokens, we use
``$q$ is full".

@@

\begin{proof} We define $\alpha$ inductively, from leaves upwards.
For a leaf $\ell$ in $\pruned$, set $\alpha_\ell = |H_\ell|$. Consider internal node $p$
with children $q_1, q_2, \ldots, q_k$. The heap $H(p)$ is formed by the union
of the $H(q_i)$ heaps, after deleting certain vertices from them, and potentially
adding some new vertices. The new vertices are potentially those that come
from other nodes that have been pruned off.
The vertices in each $q_i$ as well as $\lambda(q_i)$ are deleted.
Define $\alpha_p$ to be the number of new vertices added to $H_p$.

@@

After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\pruned_p$,
denoted $a_0, a_1, a_2, \ldots, a_k$, contributes to the charge at $q$. The charge is expressed in the following
equation. Ae use the assumption that $A_{a_i} \geq (5/4) A_{a_{i-1}}$ and hence $A_{a_i} \geq (5/4)^i A_{a_0} \geq (5/4)^i$, 
as $a_0$ is an unsupported short path and hence $A_{a_0}\geq 1$. 
$$ ({\alpha_q}/99)\sum_{a_i} 1/\sqrt{A_{a_i}} \leq ({\alpha_q}/99)\sum_{a_i} (5/4)^{i/2} \leq \alpha_q/20 $$
The total charge is $\sum_{q \in \pruned_p} \alpha_p/20 = A_p/20$. 
\end{proof}

>>>>

We prove an important lemma. The cost of heap operations corresponding 
to a buffered sub-shrub can be charged to tall support chains.

@@

There is a grand total of $100 \sum_{p \in \buff} |\loss(p)|$ tokens, which
(by the buffered property of $\buff$) is strictly more than $\sum_{p \in \buff} |p|$.
We will place all these tokens on the nodes in $\buff$, using the following process.
We reverse topologically sort (descendants before ancestors) the nodes in $\buff$.
For each $p$ in order, we will place tokens in $T_p$ on the nodes of the subtree $\buff_p$.
Take an arbitrary token in $T_p$, and find the smallest (in the ordering) descendant $q$
of $p$ that currently has strictly less than $|q|$ tokens. Place the token on $q$,
and repeat. This is repeated until all descendants $q$ have $|q|$ tokens, or
all tokens in $T_p$ are used up. When a node $q$ has $|q|$ tokens, we use
``$q$ is full".

@@

\begin{proof} We define $\alpha$ inductively, from leaves upwards.
For a leaf $\ell$ in $\pruned$, set $\alpha_\ell = |H_\ell|$. Consider an internal node $p$
with children $q_1, q_2, \ldots, q_k$. The heap $H(p)$ is formed by the union
of the $H(q_i)$ heaps, after deleting certain vertices from them, and potentially
adding some new vertices. The new vertices are potentially those that come
from other nodes that have been pruned off.
The vertices in each $q_i$ as well as $\lambda(q_i)$ are deleted.
Define $\alpha_p$ to be the number of new vertices added to $H_p$.

@@

After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\pruned_p$,
denoted $a_0, a_1, a_2, \ldots, a_k$, contributes to the charge at $q$. The charge is expressed in the following
equation. We use the assumption that $A_{a_i} \geq (5/4) A_{a_{i-1}}$ and hence $A_{a_i} \geq (5/4)^i A_{a_0} \geq (5/4)^i$, 
as $a_0$ is an unsupported short path and hence $A_{a_0}\geq 1$. 
$$ ({\alpha_q}/99)\sum_{a_i} 1/\sqrt{A_{a_i}} \leq ({\alpha_q}/99)\sum_{a_i} (5/4)^{i/2} \leq \alpha_q/20 $$
The total charge is $\sum_{q \in \pruned_p} \alpha_p/20 = A_p/20$. 
\end{proof}
